---
title: "Analyzing the Variation in Melanoma Cell Phenotype to Observe Homeostatic Changes to Cell State"
author: "Samuel Lukas, Gibran Kapadia, Arhan Rupani, Ella Bai"
subtitle: 
date: December 8, 2022
output: 
  beamer_presentation:
    theme: "Pittsburgh"
    colortheme: "orchid"
    fonttheme: "structurebold"
    slide_level: 2
classoption: "aspectratio=169"
fontsize: 11pt
urlcolor: blue
---
```{r, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(rpart)
library(partykit)
library(broom)
library(latex2exp)
library(gridExtra) #install.packages(gridExtra)
library(kableExtra) #install.packages(kableExtra)

dataset <- read_csv("STA130_Course_Project.csv")
```

## Introduction

- Wrangle and clean the dataset provided which contains the different protein and transcription levels at varying dose, time, and drug treatment
- Explore the various relationships between the MiTFg, SOX10, NGFR, AXL proteins and the 22 transcription factors
- Interpret the results and relate them to the 4 different melanoma cell phenotypes
- Form a general hypothesis about how these different factors and variables can cause different homeostatic changes that result in harmful or beneficial phenotypes

---

## Objectives

1. Observe how protein levels in an arbitrary experimental condition change over time

2. Determine the relationship between different proteins at varying experimental conditions at times

3. Create a model predicting cellular phenotypical outcomes or values/states form transcription factors

4. Decipher the results and conclude any patterns or commonalities between the analyses

---

## Statistical Methods

Throughout the analysis we will use various techniques to come to conclusions:

1. Hypothesis Testing

2. Correlation Estimation

3. Classification Modelling

---

## MiTFg with Ven: 0.316 uM dose, 0.5h vs. 120h

:::columns

::::column

- $H_0$: There is no change in the level of MiTFg protein between times when Vem is administered
- $H_a$: There is a change in the level of MiTFg protein between times when Vem is administered
- At the alpha significance level of 0.01, we have enough evidence to reject the null hypothesis and conclude there is a difference in MiTFg level at different times

::::


::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
MiTFg1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 1) %>% select(MiTFg)
MiTFg2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 1) %>% select(MiTFg)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(MiTFg1$MiTFg)
mean_t2 <- mean(MiTFg2$MiTFg)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::

:::

---

## MiTFg with Vem+Tram: 0.316uM dose, 0.5h vs. 120h

:::columns

::::column

- $H_0$: There is no change in the level of MiTFg protein between times when Vem and Tram is administered
- $H_a$: There is a change in the level of MiTFg protein between times when Vem and Tram is administered
- At the alpha significance level of 0.01, we don't have enough evidence to reject the null hypothesis

::::


::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
MiTFg1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 2) %>% select(MiTFg)
MiTFg2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 2) %>% select(MiTFg)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(MiTFg1$MiTFg)
mean_t2 <- mean(MiTFg2$MiTFg)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::

:::


---

## SOX10 with Vem: 0.316uM dose, 0.5h vs. 120h


:::columns

::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
sox1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 1) %>% select(Sox10)
sox2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 1) %>% select(Sox10)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(sox1$Sox10)
mean_t2 <- mean(sox2$Sox10)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::

::::column

- $H_0$: There is no change in the level of SOX10 protein between times when Vem is administered
- $H_a$: There is a change in the level of SOX10 protein between times when Vem is administered
- At the alpha significance level of 0.01, we don't have enough evidence to reject the null hypothesis

::::

:::

---

## SOX10 with Vem+Tram: 0.316uM dose, 0.5h vs. 120h


:::columns


::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
sox1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 2) %>% select(Sox10)
sox2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 2) %>% select(Sox10)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(sox1$Sox10)
mean_t2 <- mean(sox2$Sox10)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::


::::column

- $H_0$: There is no change in the level of SOX10 protein between times when Vem and Tram is administered
- $H_a$: There is a change in the level of SOX10 protein between times when Vem and Tram is administered
- Similarly, at the alpha significance level of 0.01, we don't have enough evidence to reject the null hypothesis

::::


:::


---

## NGFR with Vem: 0.316uM dose, 0.5h vs. 120h


:::columns

::::column

- $H_0$: There is no change in the level of NGFR protein between times when Vem is administered
- $H_a$: There is a change in the level of NGFR protein between times when Vem is administered
- At the alpha significance level of 0.01, we have overwhelming evidence to reject the null hypothesis and conclude that there is a change in NGFR level between times when Vem is administered

::::


::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
ngfr1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 1) %>% select(NGFR)
ngfr2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 1) %>% select(NGFR)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(ngfr1$NGFR)
mean_t2 <- mean(ngfr2$NGFR)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::

:::

---

## NGFR with Vem+Tram: 0.316uM dose, 0.5h vs. 120h


:::columns

::::column

- $H_0$: There is no change in the level of NGFR protein between times when Vem and Tram is administered
- $H_a$: There is a change in the level of NGFR protein between times when Vem and Tram is administered
- Similarly, at the alpha significance level of 0.01, we have overwhelming evidence to reject the null hypothesis and conclude that there is a change in NGFR level between times when Vem and Tram is administered

::::


::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
ngfr1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 2) %>% select(NGFR)
ngfr2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 2) %>% select(NGFR)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(ngfr1$NGFR)
mean_t2 <- mean(ngfr2$NGFR)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::

:::


---

## AXL with Vem: 0.316uM dose, 0.5h vs. 120h


:::columns


::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
axl1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 1) %>% select(AXL)
axl2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 1) %>% select(AXL)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(axl1$AXL)
mean_t2 <- mean(axl2$AXL)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::


::::column

- $H_0$: There is no change in the level of AXL protein between times when Vem is administered
- $H_a$: There is a change in the level of AXL protein between times when Vem is administered
- At the alpha significance level of 0.01, we have evidence to reject the null hypothesis and conclude that there is a change in AXL level between times when Vem is administered

::::


:::


---

## AXL with Vem+Tram: 0.316uM dose, 0.5h vs. 120h


:::columns


::::column

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=4, fig.align='right'}
axl1 <- dataset %>% filter(timepoint_id == 1, dose_id == 3, drug_id == 2) %>% select(AXL)
axl2 <- dataset %>% filter(timepoint_id == 7, dose_id == 3, drug_id == 2) %>% select(AXL)

set.seed(131)
N<-100000
n1 <- 75
n2 <- 75

mean_t1 <- mean(axl1$AXL)
mean_t2 <- mean(axl2$AXL)

observed_test_statistic <- mean_t1 - mean_t2
permutation_test_statistics<-1:N

x1 <- rnorm(mean=mean_t1, n=n1)
x2 <- rnorm(mean=mean_t1, n=n2)

for(i in 1:N) {
  shuffled_xs <- sample(c(x1,x2), size=n1+n2, replace=FALSE)
  tmp <- mean(shuffled_xs[1:n1]) - mean(shuffled_xs[(n1+1):(n1+n2)])
  permutation_test_statistics[i] <- tmp
}

tibble("xbar1-xbar2"=permutation_test_statistics) %>% ggplot(aes(x=`xbar1-xbar2`)) + geom_histogram(bins=50) +
xlim(-0.8,0.8) + geom_vline(xintercept=observed_test_statistic) + ggtitle(TeX(paste("Sampling Distribution of $\\bar{x}_1 - \\bar{x}_2$ for n1=75 and n2=75 if $H_0: \\; x_1=x_2$ is TRUE"))) + theme(plot.title = element_text(size = 7))

mean(abs(permutation_test_statistics)>=abs(observed_test_statistic))
```
::::


::::column

- $H_0$: There is no change in the level of AXL protein between times when Vem is administered
- $H_a$: There is a change in the level of AXL protein between times when Vem is administered
- At the alpha significance level of 0.01, we don't have evidence to reject the null hypothesis

::::


:::

---

## Correlation between MiTFg and NGFR

:::::: {.columns}

::: {.column width="45%"}

- We measure varying dosage and drug treatment at a set time
- The correlation between MiTFg and NGFR is relatively weak
- As dosage is increased, protein levels of NGFR slightly decrease in relation to MiTFg, however, when Vem+Tram is used, NGFR increases

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}

dose_vem1 <- dataset %>% filter(drug_id == 1, dose_id == 2, timepoint_id == 1)
dose_vem2 <- dataset %>% filter(drug_id == 1, dose_id == 3, timepoint_id == 1)
dose_vt1 <- dataset %>% filter(drug_id == 2, dose_id == 2, timepoint_id == 1)
dose_vt2 <- dataset %>% filter(drug_id == 2, dose_id == 3, timepoint_id == 1)

cor1 <- cor(dose_vem1$MiTFg, dose_vem1$NGFR)
cor2 <- cor(dose_vem2$MiTFg, dose_vem2$NGFR)
cor3 <- cor(dose_vt1$MiTFg, dose_vt1$NGFR)
cor4 <- cor(dose_vt2$MiTFg, dose_vt2$NGFR)

cor_c <- c(cor1, cor2, cor3, cor4)

cors <- matrix(cor_c, nrow = 2, ncol = 2)
colnames(cors) <- c("0.1 uM", "0.316uM")
rownames(cors) <- c("Vem", "Vem + Tram")


cors %>% kbl() %>% kable_minimal(full_width = F)


```

::: 


::: {.column width="55%"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}

dose_vem1 %>% ggplot(aes(x=MiTFg, y=NGFR)) + geom_point() + labs(title="Association of NGFR vs. MiTFG with 0.1uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot1

dose_vem2 %>% ggplot(aes(x=MiTFg, y=NGFR)) + geom_point() + labs(title="Association of NGFR vs. MiTFG with 0.316uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot2

dose_vt1 %>% ggplot(aes(x=MiTFg, y=NGFR)) + geom_point() + labs(title="Association of NGFR vs. MiTFG with 0.1uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot3

dose_vt2 %>% ggplot(aes(x=MiTFg, y=NGFR)) + geom_point() + labs(title="Association of NGFR vs. MiTFG with 0.316uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot4

grid.arrange(plot1, plot2, plot3, plot4)

```



:::

::::::

---


## Correlation between MiTFg and SOX10

:::::: {.columns}

::: {.column width="45%"}

- The relationship between MiTFg and SOX10 is relatively stronger
- Correlation seems to become weaker for Vem but stays the same for Vem and Tram
- The level of SOX10 decreases as dosage of Vem is increased in relation to MiTFg. When Vem+Tram is used, protein level stays the same

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}

cor1 <- cor(dose_vem1$MiTFg, dose_vem1$Sox10)
cor2 <- cor(dose_vem2$MiTFg, dose_vem2$Sox10)
cor3 <- cor(dose_vt1$MiTFg, dose_vt1$Sox10)
cor4 <- cor(dose_vt2$MiTFg, dose_vt2$Sox10)

cor_c <- c(cor1, cor2, cor3, cor4)

cors <- matrix(cor_c, nrow = 2, ncol = 2)
colnames(cors) <- c("0.1 uM", "0.316uM")
rownames(cors) <- c("Vem", "Vem + Tram")

cors %>% kbl() %>% kable_minimal(full_width = F)


```

:::


::: {.column width="55%"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dose_vem1 %>% ggplot(aes(x=MiTFg, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. MiTFG with 0.1uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot1

dose_vem2 %>% ggplot(aes(x=MiTFg, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. MiTFG with 0.316uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot2

dose_vt1 %>% ggplot(aes(x=MiTFg, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. MiTFG with 0.1uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot3

dose_vt2 %>% ggplot(aes(x=MiTFg, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. MiTFG with 0.316uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot4

grid.arrange(plot1, plot2, plot3, plot4)
```



:::

::::::

---


## Correlation between MiTFg and AXL

:::::: {.columns}

::: {.column width="45%"}

- As shown by the table, there seems to be no correlation between MiTFg and AXL
- This means that they would not be good predictors of each other
- Protein level of AXL roughly stays the same in relation to MiTFg when dosage is increased for both drug treatments

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}

cor1 <- cor(dose_vem1$MiTFg, dose_vem1$AXL)
cor2 <- cor(dose_vem2$MiTFg, dose_vem2$AXL)
cor3 <- cor(dose_vt1$MiTFg, dose_vt1$AXL)
cor4 <- cor(dose_vt2$MiTFg, dose_vt2$AXL)

cor_c <- c(cor1, cor2, cor3, cor4)

cors <- matrix(cor_c, nrow = 2, ncol = 2)
colnames(cors) <- c("0.1 uM", "0.316uM")
rownames(cors) <- c("Vem", "Vem + Tram")

cors %>% kbl() %>% kable_minimal(full_width = F)


```

:::


::: {.column width="55%"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dose_vem1 %>% ggplot(aes(x=MiTFg, y=AXL)) + geom_point() + labs(title="Association of AXL vs. MiTFG with 0.1uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot1

dose_vem2 %>% ggplot(aes(x=MiTFg, y=AXL)) + geom_point() + labs(title="Association of AXL vs. MiTFG with 0.316uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot2

dose_vt1 %>% ggplot(aes(x=MiTFg, y=AXL)) + geom_point() + labs(title="Association of AXL vs. MiTFG with 0.1uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot3

dose_vt2 %>% ggplot(aes(x=MiTFg, y=AXL)) + geom_point() + labs(title="Association of AXL vs. MiTFG with 0.316uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot4

grid.arrange(plot1, plot2, plot3, plot4)
```



:::

::::::

---


## Correlation between NGFR and SOX10

:::::: {.columns}

::: {.column width="45%"}

- The relationship between NGFR and SOX10 is relatively weak
- The $r$ coefficient increases with dose for both drug treatments
- As dosage of both Vem and Vem+Tram is increased, the level of SOX10 increases in relation to NGFR 

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}

cor1 <- cor(dose_vem1$NGFR, dose_vem1$Sox10)
cor2 <- cor(dose_vem2$NGFR, dose_vem2$Sox10)
cor3 <- cor(dose_vt1$NGFR, dose_vt1$Sox10)
cor4 <- cor(dose_vt2$NGFR, dose_vt2$Sox10)

cor_c <- c(cor1, cor2, cor3, cor4)

cors <- matrix(cor_c, nrow = 2, ncol = 2)
colnames(cors) <- c("0.1 uM", "0.316uM")
rownames(cors) <- c("Vem", "Vem + Tram")


cors %>% kbl() %>% kable_minimal(full_width = F)


```

:::


::: {.column width="55%"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dose_vem1 %>% ggplot(aes(x=NGFR, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. NGFR with 0.1uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot1

dose_vem2 %>% ggplot(aes(x=NGFR, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. NGFR with 0.316uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot2

dose_vt1 %>% ggplot(aes(x=NGFR, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. NGFR with 0.1uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot3

dose_vt2 %>% ggplot(aes(x=NGFR, y=Sox10)) + geom_point() + labs(title="Association of SOX10 vs. NGFR with 0.316uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot4

grid.arrange(plot1, plot2, plot3, plot4)
```



:::

::::::


---

## Correlation between NGFR and AXL

:::::: {.columns}

::: {.column width="45%"}

- The relationship between NGFR and AXL is relatively weak
- The $r$ coefficient decreases with dose for both drug treatments
- The level of AXL decreases in relation to NGFR as dosage of Vem and Vem+Tram is increased

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}

cor1 <- cor(dose_vem1$NGFR, dose_vem1$AXL)
cor2 <- cor(dose_vem2$NGFR, dose_vem2$AXL)
cor3 <- cor(dose_vt1$NGFR, dose_vt1$AXL)
cor4 <- cor(dose_vt2$NGFR, dose_vt2$AXL)

cor_c <- c(cor1, cor2, cor3, cor4)

cors <- matrix(cor_c, nrow = 2, ncol = 2)
colnames(cors) <- c("0.1 uM", "0.316uM")
rownames(cors) <- c("Vem", "Vem + Tram")


cors %>% kbl() %>% kable_minimal(full_width = F)


```

:::


::: {.column width="55%"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dose_vem1 %>% ggplot(aes(x=NGFR, y=AXL)) + geom_point() + labs(title="Association of AXL vs. NGFR with 0.1uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot1

dose_vem2 %>% ggplot(aes(x=NGFR, y=AXL)) + geom_point() + labs(title="Association of AXL vs. NGFR with 0.316uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot2

dose_vt1 %>% ggplot(aes(x=NGFR, y=AXL)) + geom_point() + labs(title="Association of AXL vs. NGFR with 0.1uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot3

dose_vt2 %>% ggplot(aes(x=NGFR, y=AXL)) + geom_point() + labs(title="Association of AXL vs. NGFR with 0.316uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot4

grid.arrange(plot1, plot2, plot3, plot4)
```



:::

::::::


---

## Correlation between SOX10 and AXL

:::::: {.columns}

::: {.column width="45%"}

- As shown by the table, there seems to be no correlation between SOX10 and AXL
- This means that they would not be good predictors of each other
- Protein level of AXL roughly stays the same in relation to SOX10 when dosage is increased for both drug treatments

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=2}

cor1 <- cor(dose_vem1$Sox10, dose_vem1$AXL)
cor2 <- cor(dose_vem2$Sox10, dose_vem2$AXL)
cor3 <- cor(dose_vt1$Sox10, dose_vt1$AXL)
cor4 <- cor(dose_vt2$Sox10, dose_vt2$AXL)

cor_c <- c(cor1, cor2, cor3, cor4)

cors <- matrix(cor_c, nrow = 2, ncol = 2)
colnames(cors) <- c("0.1 uM", "0.316uM")
rownames(cors) <- c("Vem", "Vem + Tram")


cors %>% kbl() %>% kable_minimal(full_width = F)


```

:::


::: {.column width="55%"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dose_vem1 %>% ggplot(aes(x=Sox10, y=AXL)) + geom_point() + labs(title="Association of AXL vs. SOX10 with 0.1uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot1

dose_vem2 %>% ggplot(aes(x=Sox10, y=AXL)) + geom_point() + labs(title="Association of AXL vs. SOX10 with 0.316uM of Vem") + theme(plot.title = element_text(size = 12)) -> plot2

dose_vt1 %>% ggplot(aes(x=Sox10, y=AXL)) + geom_point() + labs(title="Association of AXL vs. SOX10 with 0.1uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot3

dose_vt2 %>% ggplot(aes(x=Sox10, y=AXL)) + geom_point() + labs(title="Association of AXL vs. SOX10 with 0.316uM of Vem+Tram") + theme(plot.title = element_text(size = 12)) -> plot4

grid.arrange(plot1, plot2, plot3, plot4)
```



:::

::::::


---

## Predicting Phenotype through Transcription Factors














```{r, message=FALSE, size="small", echo=FALSE, fig.width=7, fig.height=3}

new_data <- dataset %>% mutate(
  phenotype = case_when(
    MiTFg <= mean(dataset$MiTFg) &
    NGFR > mean(dataset$NGFR) &
    Sox10 > mean(dataset$Sox10) &
    AXL > mean(dataset$AXL) ~ "Neural crest-like",
    MiTFg > mean(dataset$MiTFg) &
    NGFR > mean(dataset$NGFR) &
    Sox10 > mean(dataset$Sox10) &
    AXL <= mean(dataset$AXL) ~ "Transitory",
    MiTFg > mean(dataset$MiTFg) &
    NGFR > mean(dataset$NGFR) &
    Sox10 <= mean(dataset$Sox10) &
    AXL <= mean(dataset$AXL) ~ "Melanocytic",
    MiTFg <= mean(dataset$MiTFg) &
    NGFR <= mean(dataset$NGFR) &
    Sox10 <= mean(dataset$Sox10) &
    AXL > mean(dataset$AXL) ~ "Undifferentiated",
  )) %>% filter(!is.na(phenotype)) %>% mutate(id = row_number())

n <- nrow(new_data)
train_ids <- sample(1:n, size=round(0.8*n))
train <- new_data %>% filter(id %in% train_ids)
y_train <- train$phenotype
test <- new_data %>% filter(!(id %in% train_ids))
y_test <- test$phenotype

tree <- rpart(phenotype ~ Phospho_c_Fos+Phospho_c_Jun+Phospho_ATF2+Phospho_Fra1+c_Fos+c_Jun+Fra1+JunD+ATF2+JunB+Fra2+ATF4+Phospho_ATF4+Phospho_Erk1+Phospho_ATF1+ATF6+Phospho_S6+ATF3+ATF5+Phospho_p38+Ki_67+NF_kappaB, data=train)
tree %>% as.party() %>% plot(type="simple", gp=gpar(cex=0.5), ep_args = list(justmin=15),
  tp_args = list(id=FALSE, FUN = function(info) list(
    format(info$prediction), format(paste("n =",info$n)), 
    format(paste(round(info$err,1),"% err",sep="")))))

```


### Based on solely transcription factors, cells are most likely to be either Transitory or Undifferentiated

---

## Predicting Phenotype Continued


### Along with our previous observations, we can start to predict certain phenotypes

- The terminal nodes have high error so they may not be the best predictors
- However, certain transcription factors such as PhosphoS6, ATF4, and Ki67 may lead to an undifferentiated melanoma cell
- Similarly, transcription factors such as Ki67, ATF4, and Fra2 may lead to a transitory melanoma cell



```{r, echo=FALSE, message=FALSE, warning=FALSE}


phen <- matrix(c("LOW", "LOW", "HIGH", "HIGH", "LOW", "HIGH", "HIGH", "HIGH", "LOW", "HIGH", "HIGH", "LOW", "HIGH", "HIGH", "LOW", "LOW"), ncol = 4, nrow = 4)

colnames(phen) <- c("MiTFg", "NGFR", "SOX10", "AXL")

rownames(phen) <- c("Undifferentiated", "Neural crest-like", "Transitory", "Melanocytic")

phen %>% kbl() %>% kable_styling()



```




---

## Meta-analysis

### Protein level in varying time and drug treatment: 
- As time increases, levels of NGFR seems to increase while SOX10 stays roughly the same, and both AXL and MiTFg seem to decrease
- When Vem is used, the protein levels of MiTFg, SOX10, and AXL seems to decrease the most in comparison to Vem and Tram

### Protein Correlation
- We want low dosage since the correlation between MiTFg and SOX10 is the strongest
- From previous analysis, MiTFg decrease with time and a high $r$ value suggests a significant decrease in SOX10 which is desirable 
- A low correlation would be a high risk in the case of NGFR and AXL, hence, we want a high correlation and thus low dosage in order to minimize risk

---

## Meta-analysis continued

To encourage healthy homeostasis, we want to avoid certain transcription factors and instead encourage the presence of factors that result in undifferentiated cells

### Transcription Factors


| Phenotype                          	| Transcription Factors                                                                                                  	|
|------------------------------------	|------------------------------------------------------------------------------------------------------------------------	|
| - Undifferentiated                 	| - PhosphoS6, ATF4, and Ki_67 may lead to low levels of MiTFg, NGFR, and SOX10 and high levels of AXL                   	|
| - Neural Crest-Like                	| - Ki_67, Fra2, and c_Jun can be used to predict low levels of MiTFg with high levels of NGFR, SOX10, and AXL           	|
| - Transitory                       	| - Ki_67, ATF4, and Fra2 may lead to high levels of MiTFg, NGFR, and SOX10 and low levels of AXL                        	|
| - Melanocytic                      	| - Ki_67, Fra2, ATF4, or c_Fos could potentially lead to high levels of MiTFg and NGFR, and low levels of SOX10 and AXL 	|


---

## Conclusion

### To keep the cell as undifferentiated and healthy, MiTFg, SOX10, and NGFR must be low while AXL is high, hence:
- We should use Vem with a relatively low dosage with an average amount of time to keep protein level around the same
- Since both SOX10 and MiTFg have no correlation in terms of dosage with AXL, only the dosage between AXL and NGFR matters
- We want to avoid treatments that increase levels of Ki_67, Fra2, ATF4, or c_Fos

### Limitations and Assumptions:
- We conduct these statistical test assuming that only dosage, time, and drug affects protein level and no other confounding variable has underlying affects on the data
- Data and results can not be generalized or extrapolated to other carcinogenic cells since data collected is only about potentially carcinogenic melanoma cells

